\ \\ \LARGE\bf Speculative Evaluation in Particle Swarm Optimization \thanks{Matthew Gardner, Andrew McNabb, and Kevin Seppi are with the Department of Computer Science, Brigham Young University, 3361 TMCB, Provo, UT 84602 (phone: 801-422-8717; email: \{mjg82,a,k\}@cs.byu.edu).}

### 2010-05-28 honors_thesis/journal_paper/paper.tex
A few minor fixes to the paper; I changed the title

\title{\ \\ \LARGE\bf Speculative Evaluation in Particle Swarm Optimization%
\thanks{Matthew Gardner, Andrew McNabb, and Kevin Seppi are with the Department
of Computer Science, Brigham Young University, 3361 TMCB, Provo, UT 84602
(phone: 801-422-8717; email: \{mjg82,a,k\}@cs.byu.edu).}%
}

@@

As can be seen in \fig{schwefel2}, almost all of the speculative algorithms
perform worse than PSO Complete, with Many Iterations performing comparably to
PSO.  However, one of the speculative algorithms did perform better than PSO,
Pick Best Pruned.  It seems that the larger swarm size available because of
pruning was the cause of the improved performance.

@@

With Griewank, the premature convergence problems inherent in picking the best
child are exacerbated when speculating several iterations ahead.  When Many
Iterations finds the optimum, it finds it quicker than any other method we
tried, on average four times faster than standard PSO.  However, it also gets
stuck and fails to find the optimum more than any other method.  The results
are shown in \fig{griewank-manyiters}.

@@

Now that we have seen the effect that the various algorithms have on speed of
convergence and success rates, we show a result for using 800 processors, where
most of the algorithms have a 100\% success rate.  In this case, the best
algorithm to use is the one which finds the optimum the fastest.  As can be
seen in \fig{griewank-800}, with 800 processors the best choice is Pick Best.

@@

Bohachevsky is a unimodal function best optimized with a Complete swarm.  It is
similar to Griewank in that there is a global optimum with a value of 0, and
the swarm sometimes finds it and sometimes does not.  Since we have already
demonstrated the convergence properties of the various algorithms with
Griewank, for Bohachevsky we merely present the best results.  In
\fig{bohachevsky} we can see that all of the speculative algorithms found the
optimum much quicker than PSO with a Random topology.  However, SEPSO was
slower than PSO Complete and got stuck around 15\% of the time.  Using Pick
Best, Pick Best Pruned, and Many Iterations all outperformed PSO Complete,
though Many Iterations only almost reached a 100\% success rate.

>>>>

\title{\ \\ \LARGE\bf A Speculative Approach to Parallelization in Particle
Swarm Optimization%
\thanks{Matthew Gardner, Andrew McNabb, and Kevin Seppi are with the Department
of Computer Science, Brigham Young University, 3361 TMCB, Provo, UT 84602
(phone: 801-422-8717; email: \{mjg82,a,k\}@cs.byu.edu).}%
}

@@

As can be seen in \fig{schwefel2}, almost all of the speculative algorithms
perform worse than PSO Complete, with Many Iterations performing comparably to
PSO.  However, one of the speculative algorithms did perform better than PSO:
Pick Best Pruned.  It seems that the larger swarm size available because of
pruning was the cause of the improved performance.

@@

With Griewank, the premature convergence problems inherent in picking the best
child are exacerbated when speculating several iterations ahead.  When Many
Iterations finds the optimum, it finds it quicker than any other method we
tried, on average four times faster than standard PSO.  However, it also gets
stuck and fails to find the optimum more than any other method.  The results
are shown in \fig{griewank-manyiters}.   \fig{griewank-manyiters} is also
interesting in that it highlights the trade-off between accuracy and speed in
the various algorithms at this swarm size.  The faster the algorithm finds the
optimum, the less likely it is to be successful.

@@

Now that we have seen the effect that the various algorithms have on speed of
convergence and success rates, we show a result for using 800 processors, where
most of the algorithms have a 100\% success rate.  In this case, the best
algorithm to use is the one which finds the optimum the fastest.  As can be
seen in \fig{griewank-800}, with 800 processors the best choice is Pick Best,
which finds the optimum on average three times as fast as standard PSO.

@@

Bohachevsky is a unimodal function best optimized with a Complete swarm.  It is
similar to Griewank in that there is a global optimum with a value of 0, and
the swarm sometimes finds it and sometimes does not.  Since we have already
demonstrated the convergence properties of the various algorithms with
Griewank, for Bohachevsky we merely present the best results.  In
\fig{bohachevsky} we can see that all of the speculative algorithms found the
optimum much quicker than PSO with a Random topology.  However, SEPSO was
slower than PSO Complete and got stuck around 15\% of the time.  Pick Best,
Pick Best Pruned, and Many Iterations all outperformed PSO Complete, though
Many Iterations only almost reached a 100\% success rate.
